Capstone Movielens Report

### 2025-05-10 reports/capstone-movielens-report.draft4.tex
PDF doc updated.

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-09}

@@

Below, we provide the most important code of the training our model
using the \texttt{5-Fold\ Cross\ Validation} method:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Preconfiguration:} Preliminary determination of the optimal
  range of \(\lambda\) values for the \texttt{5-Fold\ Cross\ Validation}
  samples;
\item
  \textbf{Fine-tuning:} figuring out the value of \(\lambda\) that
  minimizes the model's RMSE.
\end{enumerate}

@@

\begin{noteblock}
The functions described below are defined in the \href{}{Model Tuning}
section of the \href{}{common-helper.functions.R} script.

@@

\subparagraph{\texorpdfstring{\texttt{tune.model\_param}
Function}{tune.model\_param Function}}\label{tune.model_param-function}

@@

@@

  \FunctionTok{put\_log}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{param\_values:"}\NormalTok{)}
  \FunctionTok{put}\NormalTok{(param\_values)}

  \ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{n) \{}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Iteration \%1"}\NormalTok{, i)}
\NormalTok{    prm\_val }\OtherTok{\textless{}{-}}\NormalTok{ param\_values[i]}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{prm\_val: \%1"}\NormalTok{, prm\_val)}
\NormalTok{    param\_vals\_tmp[i] }\OtherTok{\textless{}{-}}\NormalTok{ prm\_val}

    \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{param\_vals\_tmp[\%1]: \%2"}\NormalTok{, i, param\_vals\_tmp[i])}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{param\_vals\_tmp length: \%1"}\NormalTok{, }\FunctionTok{length}\NormalTok{(param\_vals\_tmp))}
    \FunctionTok{put}\NormalTok{(param\_vals\_tmp)}

\NormalTok{    RMSE\_tmp }\OtherTok{\textless{}{-}} \FunctionTok{fn\_tune.test.param\_value}\NormalTok{(prm\_val)}

    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{RMSE\_tmp: \%1"}\NormalTok{, RMSE\_tmp)}
\NormalTok{    RMSEs\_tmp[i] }\OtherTok{\textless{}{-}}\NormalTok{ RMSE\_tmp}

    \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{RMSEs\_tmp[\%1]: \%2"}\NormalTok{, i, RMSEs\_tmp[i])}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{RMSEs\_tmp length: \%1"}\NormalTok{, }\FunctionTok{length}\NormalTok{(RMSEs\_tmp))}
    \FunctionTok{put}\NormalTok{(RMSEs\_tmp)}

    \FunctionTok{plot}\NormalTok{(param\_vals\_tmp[RMSEs\_tmp }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{], RMSEs\_tmp[RMSEs\_tmp }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{])}
    \CommentTok{\# browser()}

@@

      \ControlFlowTok{if}\NormalTok{ (i }\SpecialCharTok{\textgreater{}}\NormalTok{ i\_max.beyond\_RMSE\_min) \{}
        \FunctionTok{warning}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Operation is breaked (after \textasciigrave{}RSME\textasciigrave{} reached its minimum) on the following step: "}\NormalTok{, i)}
        \CommentTok{\# browser()}
        \ControlFlowTok{break}
\NormalTok{      \}}
      \CommentTok{\# browser()}
      \ControlFlowTok{next}
\NormalTok{    \}}

@@

    \ControlFlowTok{if}\NormalTok{ (break.if\_min) \{}
\NormalTok{      i\_max.beyond\_RMSE\_min }\OtherTok{\textless{}{-}}\NormalTok{ i }\SpecialCharTok{+}\NormalTok{ steps.beyond\_min}
\NormalTok{    \}}
    \CommentTok{\# browser()}
\NormalTok{  \}}

@@

  \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Completed with RMSEs\_tmp length: \%1"}\NormalTok{, }\FunctionTok{length}\NormalTok{(RMSEs\_tmp))}
  \FunctionTok{list}\NormalTok{(}\AttributeTok{tuned.result =} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{RMSE =}\NormalTok{ RMSEs\_tmp,}
                                \AttributeTok{parameter.value =}\NormalTok{ param\_vals\_tmp),}
       \AttributeTok{best\_result =}\NormalTok{ param\_values.best\_result)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

@@

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-84-1.pdf}

>>>>

\title{Capstone Movielens Report}
\author{Azamat Kurbanaev}
\date{2025-05-10}

@@

Below, we provide the most significant part of the code for training our
model using the \texttt{5-Fold\ Cross\ Validation} method:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Preconfiguration:} Preliminary determination of the optimal
  range of \(\lambda\) values for the \texttt{5-Fold\ Cross\ Validation}
  samples;
\item
  \textbf{Fine-tuning:} figuring out the value of \(\lambda\) that
  minimizes the model's RMSE.
\end{enumerate}

@@

\begin{noteblock}
The full source code of the functions described below are are available
in the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L74}{Model
Tuning} section of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R}{common-helper.functions.R}
script on \emph{GitHub}.

@@

\subparagraph{\texorpdfstring{\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L116}{tune.model\_param}
Function}{tune.model\_param Function}}\label{tune.model_param-function}

@@

Below is the most significant part of the code of the
\href{https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R\#L116}{tune.model\_param}
function:

@@

  \CommentTok{\# ...}

  \ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{n) \{}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Iteration \%1"}\NormalTok{, i)}
\NormalTok{    prm\_val }\OtherTok{\textless{}{-}}\NormalTok{ param\_values[i]}
\NormalTok{    param\_vals\_tmp[i] }\OtherTok{\textless{}{-}}\NormalTok{ prm\_val}

\NormalTok{    RMSE\_tmp }\OtherTok{\textless{}{-}} \FunctionTok{fn\_tune.test.param\_value}\NormalTok{(prm\_val)}
\NormalTok{    RMSEs\_tmp[i] }\OtherTok{\textless{}{-}}\NormalTok{ RMSE\_tmp}

    \FunctionTok{plot}\NormalTok{(param\_vals\_tmp[RMSEs\_tmp], RMSEs\_tmp[RMSEs\_tmp])}

@@

      \ControlFlowTok{if}\NormalTok{ (i }\SpecialCharTok{\textgreater{}}\NormalTok{ i\_max.beyond\_RMSE\_min) \{}
        \FunctionTok{warning}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Operation is breaked (after \textasciigrave{}RSME\textasciigrave{} reached its minimum) on the following step: "}\NormalTok{, i)}
        \ControlFlowTok{break}
\NormalTok{      \}}
      \ControlFlowTok{next}
\NormalTok{    \}}

@@

    \ControlFlowTok{if}\NormalTok{ (break.if\_min) \{}
\NormalTok{      i\_max.beyond\_RMSE\_min }\OtherTok{\textless{}{-}}\NormalTok{ i }\SpecialCharTok{+}\NormalTok{ steps.beyond\_min}
\NormalTok{    \}}
\NormalTok{  \}}

@@

  \FunctionTok{list}\NormalTok{(}\AttributeTok{tuned.result =} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{RMSE =}\NormalTok{ RMSEs\_tmp,}
                                \AttributeTok{parameter.value =}\NormalTok{ param\_vals\_tmp),}
       \AttributeTok{best\_result =}\NormalTok{ param\_values.best\_result)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Parameters

\textbf{param\_values:}; \textbf{fn\_tune.test.param\_value:};
\textbf{break.if\_min = TRUE:}; \textbf{steps.beyond\_min = 2:}:.

\begin{itemize}
\tightlist
\item
  \textbf{param\_values:} A list of values to search for the value
  corresponding to the minimum value of the \texttt{RMSE} ;
\item
  \textbf{fn\_tune.test.param\_value:} A function that calculates the
  value of the \texttt{RMSE} for a given parameter value.;
\item
  \textbf{break.if\_min = TRUE:} A Boolean parameter that determines
  whether the function should terminate after completing the number of
  steps specified by the parameter \texttt{steps.beyond\_min}, after the
  minimum value of the \texttt{RMSE} has been found;
\item
  \textbf{steps.beyond\_min = 2:} (takes effect only if
  \texttt{break.if\_min} parameter is \texttt{TRUE}) Specifies the
  number of steps after finding the minimum value of the \texttt{RMSE},
  upon completion of which the function should terminate.
\end{itemize}

Details

The function searches for the parameter value corresponding to the
minimum value of the \texttt{RMSE} from the list of values specified by
the \texttt{param\_values} parameter. During execution, the function
uses a helper function specified by the
\texttt{fn\_tune.test.param\_value} parameter, which calculates the
\texttt{RMSE} value for the given parameter from the list determined by
the \texttt{param\_values} parameter.

\begin{noteblock}
Note that the function assumes that the dependence of the mean square
error on the input parameter is a monotonically decreasing function
until a minimum is reached and monotonically increasing thereafter. That
is, it is assumed to have a unique minimum in the domain of definition.

Value

The function returns a data structure containing the found value of the
input parameter \texttt{param\_values} for which the \texttt{RMSE} value
is minimal, as well as the minimum \texttt{RMSE} value itself, along
with a sequence of all calculated \texttt{RMSE} values.

@@

\subparagraph{`` Function}\label{function-3}

@@

\subparagraph{`` Function}\label{function-4}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-86-1.pdf}
